Assignment 1: Data and Descriptive

### 2023-11-02 assignment_1.qmd

library(tidyverse)
library(tinytex)
library(ggplot2)
library(readxl)
library(magrittr)
library(vtable)
library(lubridate)
library(dineq)
```

@@

>>>>

library(tidyverse)
library(tinytex)
library(ggplot2)
library(readxl)
library(magrittr)
library(vtable)
library(lubridate)
library(dineq)
library(vtable)
library(dineq)
library(dplyr)
library(huxtable)
library(lmtest)
library(plm)
library(readr) 
```

@@

## **I. Data Collection**

```{r}
#using R and the dplyr package to manipulate a dataset and calculate GDP per capita at the NUTS2 region level for the year 2010

#GDP_Per_Capita_NUTS2 <- GDP_Per_Capita %>%
  #filter(Year == 2010) %>%
  #select(Year, Population, GDP, NUTS2, GINI_NUTS2, NUTS) %>%
  #group_by(NUTS2, Year) %>%
  #mutate(GDP_NUTS2 = sum(GDP), Population_NUTS2 = sum(Population)) %>%
  #ungroup() %>%
  #distinct(NUTS, NUTS2, Year, GDP_NUTS2, Population_NUTS2, GINI_NUTS2) %>%
  #mutate(GDP_capita_NUTS2 = (GDP_NUTS2 * 1000000)/Population_NUTS2)

```

```{r}

#using R and the dplyr package to load and manipulate three different datasets related to demographics, transport infrastructure, and education at the NUTS2 regional level. Below is a breakdown of the steps and transformations you are performing on each dataset:

#Demographics_NUTS2 <- read.csv('demo_r_mlifexp__custom_3767848_linear.csv') %>%
  #select(geo, TIME_PERIOD, OBS_VALUE) %>%
  #mutate(Life_expectency = OBS_VALUE, Year = TIME_PERIOD, NUTS2 = geo) %>%
  #select(NUTS2, Year, Life_expectency)

#Trans_infrastructure_NUTS2 <- read.csv('tran_r_net__custom_3761027_linear.csv') %>%
  #select(geo, TIME_PERIOD, OBS_VALUE, tra_infr) %>%
  #mutate(Number_of_MWAY_and_RL = OBS_VALUE, Year = TIME_PERIOD, NUTS2 = geo, #MWAY_or_RL = tra_infr) %>%
  #select(NUTS2, Year, MWAY_or_RL, Number_of_MWAY_and_RL)

#Education_NUTS2 <- read.csv('edat_lfse_04__custom_3767870_linear.csv') %>%
  #filter(age == "Y25-64") %>%
  #select(geo, TIME_PERIOD, OBS_VALUE) %>%
  #mutate(Tertiary_educ = OBS_VALUE, Year = TIME_PERIOD, NUTS2 = geo) %>%
  #select(NUTS2, Year, Tertiary_educ)
```

```{r}
#Integrating different datasets into a comprehensive dataframe named Inequity_Data, and subsequently creating separate dataframes for specific countries based on their NUTS (Nomenclature of Units for Territorial Statistics) codes and the year 2010. Here's a breakdown of the code:

Inequity_Data <- GDP_Per_Capita_NUTS2 %>%
  left_join(Education_NUTS2, by = c("Year", "NUTS2")) %>%
  left_join(Demographics_NUTS2, by = c("Year", "NUTS2")) %>%
  left_join(Trans_infrastructure_NUTS2, by = c("Year", "NUTS2")) %>%
  mutate(log_GDP_capita_NUTS2 = log(GDP_capita_NUTS2)) %>%
  filter(Year == 2010 & GINI_NUTS2 > 0)

AT2010 <- Inequity_Data %>%
  filter(NUTS == "PT")
BE2010 <- Inequity_Data %>%
  filter(NUTS == "FR")
BG2010 <- Inequity_Data %>%
  filter(NUTS == "HU")
HR2010 <- Inequity_Data %>%
  filter(NUTS == "SK")
IT2010 <- Inequity_Data %>%
  filter(NUTS == "DK")
```

## **II Model specification**

**Simple Linear Regression**

```{r}
#creating linear regression models for seven different datasets, each corresponding to a different country or the entire dataset, to analyze the relationship between regional inequality (measured by the GINI_NUTS2 variable) and per capita GDP (measured by the GDP_capita_NUTS2 variable) for the year 2010.

#lm1 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = PT2010)
#lm2 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = FR2010)
#lm3 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = HU2010)
#lm4 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = SK2010)
#lm5 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = DK2010)
#lm6 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = Inequity_Data)
```

Table XXX

```{r}
#using the huxreg() function to create a table summarizing the results of several linear regression models. The huxreg() function is typically used with the huxtable package in R, which is a tool for creating LaTeX and HTML tables from R objects.

#huxreg(list("PT" = lm1, "FR" = lm2,  "HU" = lm3, "SK" = lm4, "DK" = lm5, "CC" = lm6), statistics = c(N = "nobs", R2 = "r.squared"), coefs = c(GDP_capita_NUTS2 = "GDP_capita_NUTS2", const. = "(Intercept)"), note = "Note: {stars} T statistics in brackets.")
```

PLOTTING

```{r}
#creating a scatter plot of the GINI_NUTS2 against GDP_capita_NUTS2 using data from Inequity_Data. In addition to the scatter plot, a regression line (calculated from a previous simple linear regression model, presumably) is added to the plot.

#| label: fig-17
#| fig-cap: "Gini & GDP per capita CC"
#Inequity_Data %>%
  #ggplot(aes(x = GDP_capita_NUTS2, y = GINI_NUTS2)) +
  #geom_point(color = "maroon", alpha = 0.7) +
  #geom_abline(intercept = b0_GDP_on_GINI_IT, slope = b1_GDP_on_GINI_IT, color = #"blue", alpha = 0.5) +
  #labs(x = "GDP per Capita", y = "GINI") +
  #theme_bw()

```

## **III. Residuales / model interpetation?**

```{r}
#predicting the GINI coefficient based on GDP per capita for all the regions in Inequity_Data using the lm_GDP_on_GINI model, and then adding these predicted values and residuals to the Inequity_Data dataframe.

# Fit a simple linear regression model
#lm_GDP_on_GINI <- lm(GINI_NUTS2 ~ GDP_capita_NUTS2, data = Inequity_Data)

# Predict the GINI coefficient and add predictions and residuals to the dataframe
#Inequity_Data$Predicted <- predict(lm_GDP_on_GINI)
#Inequity_Data$Residuals <- residuals(lm_GDP_on_GINI)

#Inequity_Data$Predicted <- predict(lm_GDP_on_GINI)
#Inequity_Data$Residuals <- residuals(lm_GDP_on_GINI)
```

```{r}
#| label: fig-18
#| fig-cap: "Gini & GDP per capita CC with residuals"

#predicting the GINI coefficient based on GDP per capita for all the regions in Inequity_Data using the lm_GDP_on_GINI model, and then adding these predicted values and residuals to the Inequity_Data dataframe.

#Inequity_Data %>%
  #distinct(NUTS2, GDP_capita_NUTS2, .keep_all = TRUE) %>%
  #ggplot(aes(x = GDP_capita_NUTS2, y = GINI_NUTS2)) +
  #geom_point(color = "maroon", alpha = 0.7) +
  #geom_segment(aes(x = GDP_capita_NUTS2, xend = GDP_capita_NUTS2, y = GINI_NUTS2, #yend = Predicted), color = "red", alpha = 0.4) +
                 #geom_abline(intercept = b0_GDP_on_GINI, slope = b1_GDP_on_GINI, color = "blue", alpha = 0.5) +
                 #labs(x = "GDP per Capita", y = "GINI") + theme_bw()
```

## Other determinants of inequity

```{r}

#defines a function named MRM_NUTS2 that performs multiple linear regression analyses on a given dataset (mlrdata). This function creates five linear models, each with different sets of predictors, and then displays the summary of these models in a nicely formatted table using the huxreg function from the huxtable package.

#MRM_NUTS2 <- function(mlrdata){
  #lm8 = lm(GINI_NUTS2 ~ Tertiary_educ, mlrdata)
  #lm9 = lm(GINI_NUTS2 ~ Number_of_MWAY_and_RL, mlrdata)
  #lm10 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2 + Tertiary_educ + Number_of_MWAY_and_RL, #mlrdata)
  #lm11 = lm(GINI_NUTS2 ~ Life_expectency, mlrdata)
  #lm12 = lm(GINI_NUTS2 ~ GDP_capita_NUTS2 + Tertiary_educ + Life_expectency, #mlrdata)

  #huxreg(
    #list("Model 1" = lm8, "Model 2" = lm9, "Model 3" = lm10, "Model 4" = lm11, "Model 5" = lm12), statistics = c(N = "nobs", R2 = "r.squared"), coefs = c(Higher.educ = "Tertiary_educ","Number_of_MWAY_and_RL", Life.expec = "Life_expectency", const. = "(Intercept)"),note = "Note: {stars} T statistics in brackets.")

```

### MLR by country

#### Portugal

```{r}
#| label: MRL Portugal
#| tbl-cap: "MRM PT"
#MRM_NUTS2(AT2010)
```

#### France

```{r}
#| label: MLR France
#| tbl-cap: "MRM FR"
MRM_NUTS2(AT2010)
```

#### Hungary

```{r}
#| label: MRL Hungary
#| tbl-cap: "MRM AT"
MRM_NUTS2(AT2010)
```

#### Slovakia

```{r}
#| label: MRL Slovakia
#| tbl-cap: "MRM AT"
MRM_NUTS2(AT2010)
```

#### Denmark

```{r}
#| label: MRL Denmark
#| tbl-cap: "MRM AT"
MRM_NUTS2(AT2010)
```
